---
title: "Integrate gene expression datasets, divided by Batch and Donor"
author: "Raymond Louie"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(results = 'asis')
knitr::opts_chunk$set(fig.width = 25)
knitr::opts_chunk$set(fig.height = 10)
knitr::opts_chunk$set(fig.pos = '!h')
knitr::opts_knit$set(root.dir = getwd())
knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = difftime(Sys.time(), now)
      now <<- NULL
      # use options$label if you want the chunk label as well
      paste('Time for this code chunk:', as.character(res))
    }
  }})
)
```
```{r clear, echo=FALSE}
rm(list = ls())
graphics.off()
```


# Load libraries and read seurat objects

```{r}
library(Matrix)
library(dplyr)
library(Seurat)
library(ggplot2)
library(grid)
require(gridExtra)
library(data.table)
library(knitr)
library(kableExtra)

seurat_gene = readRDS("../adaptive_gene.rds")
seurat_protein = readRDS("../10_1016_j_cell_2021_02_018_adaptive_protein.rds")

```

The following steps mimic those in https://satijalab.org/seurat/articles/integration_introduction.html. For more information, please see this link

# Normalize each batch-donor dataset

Normalize each batch_donor data. Note that the original paper [Time-resolved systems immunology reveals a late juncture linked to fatal COVID-19, Cell 2021](https://pubmed.ncbi.nlm.nih.gov/33713619/) performed clustering on the gene expression data by regressing out Batch and Donor information. We thus assume these two factors contribute to unwanted technical effects to the data, and thus Seurat's integration pipeline will be used to correct for these technical differences.

```{r}
meta_df = seurat_protein@meta.data
meta_df$batch_donor = paste0(meta_df$Batch,"_",
                             meta_df$Donor)
unique_batch_donor = unique(meta_df$batch_donor)
# unique_batch_donor=unique_batch_donor[1:2]


seurat_gene.list = list()
for (ind in 1:length(unique_batch_donor)){
  curr_batch_donor = unique_batch_donor[[ind]]
  keep_cells = rownames(meta_df)[which(meta_df$batch_donor %in% curr_batch_donor)]
  curr_matrix = seurat_gene@assays$RNA@data[,keep_cells]
  curr_meta_gene = seurat_gene@meta.data[keep_cells,]
  curr_meta_protein = seurat_protein@meta.data[keep_cells,]
  curr_meta = cbind(curr_meta_gene,
                    curr_meta_protein[,setdiff(colnames(curr_meta_protein),
                                               colnames(curr_meta_gene))])
  
  curr_seurat <- CreateSeuratObject(counts = curr_matrix,
                                    meta.data =curr_meta)
  curr_seurat <- NormalizeData(curr_seurat)
  curr_seurat <- FindVariableFeatures(curr_seurat, selection.method = "vst", nfeatures = 2000)
  seurat_gene.list[[curr_batch_donor]] = curr_seurat    
}


```


# Integration

Choose donors/batch as references for integration, to reflect a diversity of severity.outcome and disease, otherwise integration takes too long.

```{r}
batch_donor_table= data.frame(table(paste0(meta_df$batch_donor,"_",
                                          meta_df$disease,"_",
                                          meta_df$severity.outcome)))
batch_donor_table= batch_donor_table[order(batch_donor_table$Freq,decreasing=TRUE),]

batch_donor_keep = c("B3_HGR0000051","B2_HGR0000069","B3_HGR0000430","B1_HGR0000071","B1_HGR0000083","B1_HGR0000136","B2_HGR0000124","B2_HGR0000418","B3_HGR0000429","B2_HGR0000144","B3_SHD5","B2_HDVO","B3_SHD1","B2_HDML","B2_SHD4")

```


```{r}

features <- SelectIntegrationFeatures(object.list = seurat_gene.list)
immune.anchors <- FindIntegrationAnchors(object.list = seurat_gene.list, anchor.features = features,
                                         reference=which(names(seurat_gene.list) %in% batch_donor_keep))
# this command creates an 'integrated' data assay
immune.combined <- IntegrateData(anchorset = immune.anchors)

```

# Integrated analysis and visualization

```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(immune.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
immune.combined <- ScaleData(immune.combined, verbose = FALSE)
immune.combined <- RunPCA(immune.combined, npcs = 30, verbose = FALSE)
immune.combined <- RunUMAP(immune.combined, reduction = "pca", dims = 1:30)


```


```{r}
Idents(immune.combined) = immune.combined$cell_type

# Visualization
p1 <- DimPlot(immune.combined, reduction = "umap", group.by = "Batch")
p2 <- DimPlot(immune.combined, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2

```


# Save integrated seurat object


```{r}
integrated_seurat_gene = immune.combined
saveRDS(integrated_seurat_gene,
        file="integrated_seurat_gene_ref.rds")

```



